Facility and Time Step Discretization Effects in Fuel Cycle Simulation

### 2015-10-02 paper.tex
edits up to page 21 in manuscript from dad

All results data, analysis tools and instructions are publicly available for
download and use at \url{http://dx.doi.org/10.6084/m9.figshare.1546775}
\cite{Carlsen2015}.

@@

One useful over-estimate of the fuel sharing inefficiency is to assume every
one of the "wasted" batches could have been given to a reactor enabling it to
operate.  Multiplying each of these cumulative batches by reactor capacity
(i.e. 450 MWe) provides a way to compare fuel sharing inefficiency with the
overall fuel outage.  This results in a cumulative offline power caused by
poor fuel sharing of roughly 2,000 $\text{GWe} \cdot \text{months}$ and 6,600
$\text{GWe} \cdot \text{months}$ for cases MI and QI respectively.  These
approximations are a significant portion of the difference between individual
and fleet cases in Figure \ref{fig:unfueled} suggesting that most of the
discrepency in cumulative outage between fleet and individual modeling is
caused by inefficient fuel sharing in the individual reactor cases.

@@

Having multiple assemblies per batch (item 1 above) will further degrade fuel
sharing efficiency.  Even reactors that are short only one batch could
potentially receive some fuel and still be unable to operate.  In order to
maintain optimal sharing, fresh fuel now needs to be sent in all-or-nothing
multi-assembly quanta with higher preference to reactors that need fewer
assemblies.  Spare fresh fuel inventory (item 2 above) will have the global
effect of increasing total outage severity; on average more batches of fresh
fuel will be sitting unused.  However, it can potentially reduce the frequency
of outages for individual reactors in some cases.

If they were to ever occur, real world fuel shortages would likely not result
in optimal fuel sharing. However, modeling these the causes of suboptimal fuel
sharing is probably best accomplished with a more direct, intentional approach
rather than as a modeling artifact as seen here in cases MI and QI.  One way to
alleviate the poor fuel sharing is to modify the individual reactor model to
lower the preference value for requests for fresh fuel depending on how many
assemblies it needs to have a full core.  The more assemblies it needs, the
lower it will set its request preference.  This will have the effect of
allowing the resource exchange resolution in Cyclus to globally prefer sending
fuel to reactors that need less to operate.

\begin{figure}[!h]
    \centering
    \includegraphics[width=1.0\columnwidth]{exp2/badshare.eps}
    \caption[Cumulative unnecessary idling fuel]{
        Fuel sharing inefficiency approximated by the cumulative number of
        fuel batches given to each reactor on each refueling period multiplied
        by the number of time steps that reactor had to delay the start of its
        next cycle (due to fuel shortage).  Fleet reactors have perfect fuel
        sharing by design and so are not included here. The large difference
        between cases MI and QI is mostly due to the \emph{inventory drawdown
        effect}. 
    }
    \label{fig:badshare}
\end{figure}

@@

The large differences seen between monthly and quarterly time step cases in
Figures \ref{fig:unfueled} and \ref{fig:badshare} are primarily caused by the
\emph{inventory drawdown effect}.  Figure \ref{fig:puinv} helps to visualize
this effect.  The black dots in the figure represent impulse flows of
available Pu out of the fuel fab and into fast reactors.  Blue lines represent
the flows into Pu inventory available for making fuel. The in-flows are not
shown for the individual reactor cases because they are somewhat messy and
obscure other information in the plots. The other colored curves show Pu inventory
available for fabricating fast reactor fuel.  When the inventory curves meet and
force down the out-flows, fuel shortages occur.  While this is most clear for
cases MF and QF, the individual reactor scenarios (cases MI and QI) also have many
points during the shortage where the outflow point lies exactly on top of the
inventory point - indicating that all available inventory is being
transferred.

Quarterly time steps have less frequent but larger impulse material flows.
The larger in-flows, however, do not compensate for the larger withdrawals
because facilities do not know about incoming inventory when they resolve
their outflow for a particular time step.  This information lag effectively
requires the floating Pu inventory to be maintained at a higher level during
the shortage.  This can be clearly seen in the case QF plot which has a higher
standing inventory than case MF during the fuel shortage years.  The case QF Pu
inventory is unable to drop below approximately 100 tonnes, which indicates
out-flows to be roughly equal to the Pu flow into inventory each time step
during those years.  This can be confirmed by the in-flow and inventory curves
touching at those points.  Case MF Pu inventory is similarly limited at about
the 30 tonne level.  As expected, the case QF withdrawals are also three times
higher case MF withdrawals exactly matching the time step duration ratio
between the two cases.

>>>>

All results data, analysis tools, and related instructions are publicly
available for download and use at
\url{http://dx.doi.org/10.6084/m9.figshare.1546775} \cite{Carlsen2015}.

@@

One useful over-estimate of the fuel sharing inefficiency is to assume every
one of the "wasted" batches could have been given to a reactor enabling it to
operate.  Multiplying each of these cumulative batches by the fast reactor power capacity
(i.e. 450 MWe) provides a way to compare fuel sharing inefficiency with the
overall fuel outage.  This results in a cumulative energy deficit caused by
poor fuel sharing of roughly 2,000 $\text{GWe} \cdot \text{months}$ and 6,600
$\text{GWe} \cdot \text{months}$ for cases MI and QI respectively.  These
approximations account for a significant portion of the difference between individual
and fleet cases in Figure \ref{fig:unfueled} suggesting that most of the
discrepency in cumulative outage between fleet and individual modeling is
caused by inefficient fuel sharing in the individual reactor cases.

@@

Having multiple assemblies per batch (item 1 above) will further degrade fuel
sharing efficiency.  Even reactors that are short only one batch could
potentially receive some fuel and still be unable to operate.  In order to
maintain optimal sharing, fresh fuel now needs to be sent in all-or-nothing
multi-assembly quanta with higher preference to reactors that need fewer
assemblies.  Spare fresh fuel inventory (item 2 above) will have the global
effect of increasing total shortage severity; on average more batches of fresh
fuel will be sitting unused.  However, it can potentially reduce the frequency
of outages for individual reactors in some cases.

If they were to ever occur, real world fuel shortages would likely not result
in optimal fuel sharing. However, modeling these the causes of suboptimal fuel
sharing is probably best accomplished with a more direct, intentional approach
rather than as a modeling artifact as seen here in cases MI and QI.  One way to
alleviate the poor fuel sharing is to modify the individual reactor model to
adjust the preference value on requests for fresh fuel depending on how many
assemblies it needs to have a full core.  The more assemblies it needs, the
lower it will set its request preference.  This will have the effect of
allowing the DRE in Cyclus to prefer sending
fuel to reactors that need less to operate.

\begin{figure}[!h]
    \centering
    \includegraphics[width=1.0\columnwidth]{exp2/badshare.eps}
    \caption[Cumulative unnecessary idling fuel]{
        Fuel sharing inefficiency approximated by the cumulative number of
        fuel batches given to each reactor on each refueling period multiplied
        by the number of time steps that reactor had to delay the start of its
        next cycle (due to fuel shortage).  Fleet reactors implicitly have perfect fuel
        sharing and so are not included here. The large difference
        between cases MI and QI is mostly due to the \emph{inventory drawdown
        effect}. 
    }
    \label{fig:badshare}
\end{figure}

@@

The large differences seen between monthly and quarterly time step cases in
Figures \ref{fig:unfueled} and \ref{fig:badshare} are primarily caused by the
\emph{inventory drawdown effect}.  Figure \ref{fig:puinv} helps to visualize
this effect.  The black dots in the figure represent impulse flows of
available Pu from fuel fabrication and into fast reactors.  Blue lines represent
the flows into Pu inventory available for making fuel. The in-flows are not
shown for the individual reactor cases because they are somewhat messy and
obscure other information in the plots. The other colored curves show Pu inventory
available for fabricating fast reactor fuel.  When the inventory curves meet and
force down the out-flows, fuel shortages occur.  While this is most clear for
cases MF and QF, the individual reactor scenarios (cases MI and QI) also have many
points during the shortage where the outflow point lies exactly on top of the
inventory point - indicating that all available inventory is being
transferred.

Quarterly time steps have less frequent but larger impulse material flows.
The larger in-flows, however, do not compensate for the larger withdrawals
because facilities do not know about incoming inventory when they resolve
their outflow for a particular time step.  This information lag effectively
requires the floating Pu inventory to be maintained at a higher level during
the shortage.  This can be clearly seen in the Figure \ref{fig:puinv} case QF plot which has a higher
standing inventory than case MF during the fuel shortage years.  The case QF Pu
inventory is unable to drop below approximately 100 tonnes, which indicates
out-flows to be roughly equal to the Pu flow into inventory each time step
during those years.  This can be confirmed by the in-flow and inventory curves
touching at those points.  Case MF Pu inventory is similarly limited at about
the 30 tonne level.  As expected, the case QF withdrawals are also three times
higher than the case MF withdrawals, exactly matching the time step duration ratio
between the two cases.
